White Dwarf Mergers on Adaptive Meshes\\ I. Methodology and Code Verification

### 2015-12-13 papers/wdmerger_I/paper.tex
More careful wording in the rotation section

For the evolution of binary systems, it is most natural to evolve the
two stars in a frame that is co-rotating at the same period as the
orbital period. Since the publication of the original code paper, \castro\ 
now has the ability to evolve systems in a rotating reference frame. 
Source terms corresponding to the Coriolis and centrifugal 
force terms are added to the momentum and energy equations. In this frame, 
the stars essentially remain stationary in their original positions due to the
centrifugal force supporting against the gravitational attraction, and
will remain this way as long as significant mass transfer does not
occur. \cite{swc:2000} demonstrated (in the context of neutron star
mergers) that conservation of angular momentum is much easier to
obtain in the rotating reference frame than in an inertial frame in
which stars advect large amounts of material around the domain. We
wish to emphasize that although it is commonly stated in the
literature that fixed-mesh codes poorly conserve angular momentum,
it is only generally true that mesh-based codes do not exactly conserve 
angular momentum when the equations are written in conservative form
for linear momentum. (See \cite{motl:2002} and \cite{byerly:2014} for examples of how 
to evolve the hydrodynamics equations for angular momentum.) 
The extent to which angular momentum conservation is violated
is a function of the resolution. When the resolution is sufficiently high, 
excellent conservation properties can result, as demonstrated in \autoref{sec:kepler}. 
In particular, at reasonable resolution for a binary orbit our code 
conserves angular momentum well enough that this no longer becomes 
a strong argument in favor of using a Lagrangian method.
We note that as the stars begin to coalesce, the rotating reference frame
will no longer provide a good approximation to the spatial motion of
the stars and then they will begin to significantly move around the
domain. This is not necessarily problematic because the most important
feature of the rotating frame is that it helps ensure that the initial
coalescence is not the result of spurious numerical loss of angular
momentum. When significant mass transfer sets in and evolution
proceeds on a dynamical timescale, the conservation properties may be
slightly worse but angular momentum conservation is also less
important.

>>>>

For the evolution of binary systems, it is most natural to evolve the
two stars in a frame that is co-rotating at the same period as the
orbital period. Since the publication of the original code paper, \castro\ 
now has the ability to evolve systems in a rotating reference frame. 
Source terms corresponding to the Coriolis and centrifugal 
force terms are added to the momentum and energy equations. In this frame, 
the stars essentially remain stationary in their original positions due to the
centrifugal force supporting against the gravitational attraction, and
will remain this way as long as significant mass transfer does not
occur. \cite{swc:2000} demonstrated (in the context of neutron star
mergers) that conservation of angular momentum is much easier to
obtain in the rotating reference frame than in an inertial frame in
which stars advect large amounts of material around the domain. We
wish to emphasize that although it is commonly stated in the
literature that fixed-mesh codes poorly conserve angular momentum,
it is only generally true that mesh-based codes do not exactly conserve 
angular momentum when the equations are written in conservative form
for linear momentum. Indeed, \cite{motl:2002} and \cite{byerly:2014} 
have evolved binary systems using the hydrodynamics equations written 
in a form that explicitly conserves angular momentum, and it is 
straightforward to convert an existing grid-based code to solve 
the system of equations that \citeauthor{byerly:2014} present.
Additionally, the extent to which angular momentum conservation is violated in our code
is a function of the resolution. When the resolution is sufficiently high, 
excellent conservation properties can result. At reasonable resolution 
for a binary orbit our code conserves angular momentum well enough 
to keep the stars stable for a large number of orbits; however, at moderate 
resolution in an inertial frame, there is a secular loss of angular 
momentum that eventually will result in a spurious merger.
We note that as the stars begin to coalesce, the rotating reference frame
will no longer provide a good approximation to the spatial motion of
the stars and then they will begin to significantly move around the
domain. This is not necessarily problematic because the most important
feature of the rotating frame is that it helps ensure that the initial
coalescence is not the result of spurious numerical loss of angular
momentum. When significant mass transfer sets in and evolution
proceeds on a dynamical timescale, the conservation properties may be
slightly worse but angular momentum conservation is also less
important.
